State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-06-04 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
small paper edits
some editing
fixed typo in LS algorithm
change to obseravtion noise covariance notation in the filter algorithm
commiting for merge
Connectivity Kernel Plot

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
  % \documentclass[twocolumn,11pt,a4paper]{article}		%<-use this for checking equation length
\documentclass[12pt]{iopart}

%<-comment out this for checking equations (author list too)
\usepackage{iopams,setstack} 
%<-comment out this for checking equations
\usepackage{graphicx}
\usepackage{algorithm, algorithmicx, algpseudocode} 
\usepackage{color} 
%This is to decrease the space in subfigure command
\usepackage[tight,scriptsize]{subfigure}
\makeatletter
\renewcommand{\subfigtopskip}{1\p@}
\renewcommand{\subfigcapskip}{0\p@}
\renewcommand{\subfigcaptopadj}{1\p@}
\renewcommand{\subfigbottomskip}{1\p@}
\renewcommand{\subfiglabelskip}{0.05em plus 0.02em minus 0.01em}
\makeatother
% \usepackage{pstool}
\renewcommand{\subcapsize}{\small}
 \usepackage{anysize}							%<-use this for checking equation length
 % \marginsize {1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
 %  \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
 % \usepackage{amsmath}
\usepackage{amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\mike}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}
\newcommand{\ken}[1]{\textsf{\emph{\textbf{\textcolor{magenta}{#1}}}}}
\begin{document}

@@

\address{$^1$ Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC 3010 Australia} \address{$^2$ Department of Automatic Control and Systems Engineering, University of Sheffield, Mappin Street, Sheffield, S1 3JD, UK} \address{$^3$ Department of Applied Physics and Applied Mathematics, Columbia University, NY, USA} \address{$^4$ Department of Systems and Control Engineering, University of Malta, Msida, MSD 1333, Malta} \address{$^5$ The Bionic Ear Institute, East Melbourne, VIC 3002
 Australia} \ead{dfreestone@bionicear.org}

@@

\subsection{Derivation of the Integro-difference Equation Representation}

@@

   		\includegraphics[angle = 90, scale=0.5]{./Graph/2D_kernel.pdf} 

@@

	\boldsymbol{\Psi}(\mathbf{r}') = T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi} (\mathbf{r}'-\mathbf{r})d\mathbf{r}},

@@

Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are  used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the predicted cross-covariance matrix of the states is used with the covariance matrix. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that  the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTS is used, otherwise the state vector should be augmented with the noise terms. The additive disturbance form reduces the dimension and the number of sigma points used in the algorithm.

@@

			\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1} 

@@

			\mathbf P_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{f})^\top+\boldsymbol \Sigma_e 

@@

			\mathbf M_{t +1}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{f})^\top 

@@

	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}' \boldsymbol{\theta}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[twocolumn,11pt,a4paper]{article}		%<-use this for checking equation length
% \documentclass[12pt]{iopart}

%<-comment out this for checking equations (author list too)
% \usepackage{iopams,setstack} 
%<-comment out this for checking equations
\usepackage{graphicx}
\usepackage{algorithm, algorithmicx, algpseudocode} 
\usepackage{color} 
%This is to decrease the space in subfigure command
\usepackage[tight,scriptsize]{subfigure}
\makeatletter
\renewcommand{\subfigtopskip}{1\p@}
\renewcommand{\subfigcapskip}{0\p@}
\renewcommand{\subfigcaptopadj}{1\p@}
\renewcommand{\subfigbottomskip}{1\p@}
\renewcommand{\subfiglabelskip}{0.05em plus 0.02em minus 0.01em}
\makeatother
% \usepackage{pstool}
\renewcommand{\subcapsize}{\small}
 \usepackage{anysize}							%<-use this for checking equation length
 % \marginsize {1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
 %  \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
  \usepackage{amsmath}
\usepackage{amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\mike}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}
\newcommand{\ken}[1]{\textsf{\emph{\textbf{\textcolor{magenta}{#1}}}}}
\begin{document}

@@

% \address{$^1$ Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC 3010 Australia} \address{$^2$ Department of Automatic Control and Systems Engineering, University of Sheffield, Mappin Street, Sheffield, S1 3JD, UK} \address{$^3$ Department of Applied Physics and Applied Mathematics, Columbia University, NY, USA} \address{$^4$ Department of Systems and Control Engineering, University of Malta, Msida, MSD 1333, Malta} \address{$^5$ The Bionic Ear Institute, East Melbourne, VIC 3002
%  Australia} \ead{dfreestone@bionicear.org}

@@

\subsection{Derivation of the Integro-Difference Equation Representation}

@@

   		\includegraphics{./Graph/Cross_section_kernel.pdf} 

@@

	\boldsymbol{\Psi}(\mathbf{r}') = T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}'-\mathbf{r})d\mathbf{r}},

@@

where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are  used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the predicted cross-covariance matrix of the states is used with the covariance matrix. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that  the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTS is used, otherwise the state vector should be augmented with the noise terms. The additive disturbance form reduces the dimension and the number of sigma points used in the algorithm.

@@

			\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\epsilon})^{-1} 

@@

			\mathbf P_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})^\top+\boldsymbol \Sigma_e 

@@

			\mathbf M_{t +1}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{b-})^\top 

@@

	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'.
